---
title: "Untitled"
author: "Shreya Thodupunuri"
date: "2023-11-12"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Loading the dataset
Pharmaceuticals <- read.csv("C:/Users/shrey/OneDrive/Documents/Pharmaceuticals.csv")
# Ensuring the file path is correct and the dataset is loaded properly.

```


```{r}

# Reading required libraries for data manipulation and clustering
library(tidyverse)
library(factoextra)
library(dplyr)
library(ggplot2)
library(cluster)
# These libraries are essential for manipulating data and performing cluster analysis.

```




**Task1**

kUse only the numerical variables (1 to 9) to cluster the 21 firms.
Justify the various choices made in conducting the cluster analysis, such as weights for different variables,the specific clustering algorithm(s) used, the number of clusters formed, and so on.
Removal of missing data helps in maintaining the accuracy of the cluster analysis.
Prior to clustering data, remove the missing data and rescale variables for comparability.
```{r}
# Removing missing data and selecting relevant variables for cluster analysis
Pharma_data <- na.omit(Pharmaceuticals) 
# Removing incomplete cases to maintain the accuracy of the cluster analysis.

```

```{r}
#Taking the quantitative variables(1-9) to cluster the 21 firms

row.names(Pharma_data)<- Pharma_data[,1]
Pharma_data1<- Pharma_data[,3:11]# Considering only numercial values i.e., 3-11 columns from csv file
head(Pharma_data1)

# Focusing on numerical variables (columns 3 to 11) as they are key for clustering.

```

```{r}


#Normalizing the data frame with scale method

Pharma_data2<-scale(Pharma_data1)
head(Pharma_data2)
# Scaling standardizes the data, making it suitable for clustering.

```

```{r}

# Determining the number of clusters using Elbow Method

fviz_nbclust(Pharma_data2, kmeans, method = "wss") 

# The Elbow Method helps identify the optimal cluster count by analyzing within-cluster sum of squares.

```

```{r}
##By seeing the above graph from Elbow method, Graph is not clear to choose k=2 or 3 or 4 or 5.
# Using Silhouette method to determine the number of clusters

fviz_nbclust(Pharma_data2, kmeans, method = "silhouette")


```


By seeing the graph from silhouette method, I can see sharp rise at k=5. 
So, considering the  silhouette method.

Silhouette method assesses how well each object lies within its cluster, aiding in determining the best number of clusters.


```{r}
# Performing K-means clustering
set.seed(64060)
k_5 <- kmeans(Pharma_data2, centers=5, nstart=25)

```
K-means clustering is performed with 5 clusters, determined by previous methods.



```{r}
# Visualizing cluster centroids and distances

k_5$centers
fviz_cluster(k_5,data = Pharma_data2) # to Visualize the clusters
k_5
distance<- dist(Pharma_data2, method = "euclidean")
fviz_dist(distance)

# I can see there are 5 clusters and the center is defined after 25 restarts which is determined in kmeans.
```
Visualization aids in understanding the distribution and separation of clusters.


```{r}

# Refitting K-means for a clearer interpretation
#K-Means Cluster Analysis- Fit the data with 5 clusters

fit<-kmeans(Pharma_data2,5)
```


```{r}

# Analyzing the mean values of each variable within each cluster

aggregate(Pharma_data2,by=list(fit$cluster),FUN=mean)

Pharma_data3<-data.frame(Pharma_data2,fit$cluster)
Pharma_data3
View(Pharma_data3)

```
Aggregate functions reveal the central tendencies of each cluster, highlighting their defining characteristics.

```{r}

# Cluster plot visualization

clusplot(Pharma_data2,fit$cluster,color = TRUE,shade = TRUE,labels = 2,lines = 0)
```

Clusplot visually represents the clustering, showing the grouping and outliers if any.



**Task 2** 


Interpret the clusters with respect to the numerical variables used in forming the clusters.

By noticing the mean values of all quantitative variables for each cluster

Cluster_1 - AGN, PHA, BAY - Suggests higher growth expectations or overvaluation.

Cluster_2 - JNJ, MRK, GSK, PFE - High Market Cap and Leverage: Indicates large, established companies.

Cluster_3 - AHM, AVE, WPI - Low Asset Turnover and Beta: Represents conservative, stable firms.

Cluster_4 - IVX, MRX, ELN, CHTT - Low Market Capital but High Revenue Growth: Reflects emerging growth companies.

Cluster_5 - ABT, NVS, AZN, LLY, BMY, WYE, SGP - Low Revenue Growth, High Asset Turnover, and Net Profit Margin: Signifies efficient, profitably run firms.





**Task 3**


Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)


For cluster 1: It  has the highest PE_Ratio and needs to be held as per the media recommendations.

For cluster 2: It has the highest market_Cap and has Good Leverage value. And they can be moderately recommended.

For cluster 3: It has lowest asset_turnover,and lowest beta. But media recommendations are highly positive.

For cluster 4: The leverage ratio is high, they are moderately recommended. 

For Cluster 5: They have lowest revenue growth, highest assest turnover and highest net profit margin.

They are  recommended to be held for longer time.





**Task 4**

Provide an appropriate name for each cluster using any or all of the variables in the dataset.


Cluster 1 - Balanced Performers: This name suggests that firms in this cluster have stable and decent financial metrics. It implies a balanced performance across various financial aspects.

Cluster 2 - Steady Growth Contenders: This name indicates that companies in this cluster demonstrate consistent growth, making them a moderate but reliable option for investment or holding. It reflects both stability and potential for growth.

Cluster 3 - Dynamic Opportunity Firms: This name implies that firms in this cluster might present varied investment opportunities, characterized by both potential growth (buy) and higher risk (sell). It suggests dynamism and variability in performance.

Cluster 4 - Stable Investment Picks: This name reflects firms with good stability and solid financial metrics, making them attractive for buying and long-term investment.

Cluster 5 - Long-term Value Holders: This name suggests that firms in this cluster are ideal for holding due to their potential to provide long-term value, likely characterized by lower but consistent revenue growth and high asset turnover.




